---
title: "| Título en español \n| Title in English\n"
author: "Name Apellidos^1^, Nombre Apellidos^2^"
csl: ecosistemas.csl
output:
  pdf_document: default
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 7.5
    highlight: null
    reference_docx: Ecosistemas_template.docx
bibliography:
- references.bib
- knitcitations.bib
---


> (1) Dirección del primer autor.

> (2) Dirección del segundo autor.

> Autor para correspondencia: Nombre Apellidos [minombre@micorreo.com]


# Abstract

Pollinator sharing can have negative consequences for species fitness with the arrival of foreign pollen. However, the costs of heterospecific pollen are not yet well understood. For this reason, we have conducted a glasshouse experiment where we try to understand how phylogenetic relatedness and the different traits of these species are involved in this process. We experimentally crossed 10 species belonging to three different families: Brassicaceae, Solanaceae and Convolvulaceae. Overall, more than 4000 crosses were done and seed set and pollen tubes were considered as proxy of effect. We found that for all species foreign pollen (50% or less) reduced seed set. Moreover, the seed set reduction is not dependent on the degree of relatedness of the pollen donor. However, the effect is governed by the degree of relatedness and the traits of the species recipient. Our results show that the outcome of heterospecific pollen deposition is determined in greater degree by the traits of the pollen recipient than the pollen donor and that certain traits such as compatibility system are crucial to understand the costs of heterospecific pollen.

# Keywords

> keyword1; keyword2; keyword3; keyword4

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

# Introduction


*Paragraph 1*

Plants with ecological generalization receives a wide range of floral visitors (Refs), these plant species tend to co-occur with other species and have floral visitors in common. This potential pollinators can deposit conspecific pollen and heterospecific pollen and cause conspecific pollen loss. All this features are important to understand the effectivenes of pollinators and the reproductive biology of the plant. 

*Paragraph 2*

*Paragraph 3*

*Paragraph 4*




# Methods

The study was conducted in a glasshouse at University of New England (Armidale, Australia) from November 2017 to March 2018. Rooms were temperature controlled depending on the requirements of the species with day and night temperature differences. The species selected (Table 1) belonged to three different families, Solanaceae, Brassicaceae and Convolvulaceae. The criteria of species/family selection was based on close/distant related species (see phylogenetic tree for relatedness fig 1), heterogeneous traits, low structural flower complexity and fast life cycle. For the purpose of the experiment all the species where consider as pollen recipient and as pollen donor (see interaction matrix, fig 2). Species were watered once or twice per day and fertilized weekly (NPK 23: 3.95: 14).

**Hand-pollination**

To test heterospecific pollen effect the standard 50-50% pollen mix was performed (Ref) and 100% heterospecific pollen. Furthermore, hand cross pollination, hand self pollination, apomixis (emsaculated bagged flowers) and natural selfing (non-emasculated bagged flowers) was tested.


Flowers were emasculated the day prior anthesis and hand pollinated next day with a toothpick. Had-pollination was realized with 3-4 gentle touches on the surface of the stigma. The mixes of pollen were performed based on the pollen counts on an eppendorf. Pollen on anthers was counted through Neubaeur chamber.


Treatments: 50-50% and 100%

hand self pollination

hand cross pollination (Pollen from different individual)

Apomixis

Natural selfing








## Análisis de datos

```{r read_data, eval = FALSE, echo = FALSE}
dataset <- read.csv("mydata.csv")
```

Ajustamos un modelo lineal:

$$
y_{i} = \alpha + \beta*x_{i} 
$$

```{r model, echo = FALSE, eval = FALSE}
model <- lm(y ~ x)
```

Utilizamos R `r citep(citation())` y Rmarkdown `r citep(list(citation("knitr"), citation("rmarkdown")))` para todos nuestros análisis. Para ajustar los modelos mixtos utilizamos `lme4` `r citep(citation("lme4"))`.




# Results

Esta sección está dividida en subsecciones.


## Subsección 1

Los árboles de la parcela A fueron más altos que en la parcela B (altura media: `r mean(25, 31, 28)` vs `r mean(13, 19, 16)` m). Y muchos más resultados que se actualizan dinámicamente.


## Subsección 2

Texto.


## Subsección 3

Texto.



# Discusión

Texto.



# Conclusiones

Texto.


# Agradecimientos

Agradecimientos.





###### References

```{r write_citations, cache=FALSE, include=FALSE}
write.bibtex(file="knitcitations.bib")
```

<div id = "refs"></div>














###### TABLA 1

**Tabla 1**. Cada tabla debe aportar su correspondiente encabezamiento explicativo. En los Artículos de investigación, de revisión y Comunicaciones breves se aportarán los encabezamientos tanto en castellano como en inglés, en letra Arial 10 y en página independiente. Es importante que sean simples y que no superen el ancho una página DIN A4 vertical. Los originales se deben aportar en formato tabla y no en formato imagen. 

**Table 1**. Table heading in English. 

```{r Tabla1, results='asis', echo=FALSE, cache=FALSE}
library("knitr")
kable(head(iris))
```





###### PIES DE FIGURA

**Figura 1**. Pie de figura 1. 

**Figura 2**. Pie de figura 2.





###### FIGURE LEGENDS

**Figure 1**. Figure caption.

**Figure 2**. Figure caption.








