---
title: Working title, The impact of heterospecific pollen on plant reproductive success is mediated by phylogenetic distance and floral reproductive traits 
author:
- affilnum: '1,2'
  name: Jose B. Lanuza, Ignasi Bartomeus, Tya-Linn Ashman, Romina Rader
- affilnum: 3
  email: barragansljose@gmail.com
  name: Jose
affiliation:
- affil: US Environmental Protection Agency, Office of Research and Development, National
    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
    27 Tarzwell Drive  Narragansett, RI, 02882, USA
  affilnum: 1
- affil: Big Name University, Department of R, City, BN, 01020, USA
  affilnum: 2
- affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain
  affilnum: 3
output:
  word_document: null
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    includes:
      in_header: header.tex
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
  html_document: null
capsize: normalsize
classoption: a4paper
csl: ecology.csl
documentclass: article
fontsize: 11pt
footerdate: yes
geometry: margin=1in
linenumbers: yes
bibliography:
- references.bib
- knitcitations.bib
spacing: doublespacing
abstract: no
---


```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")


```


```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```


\singlespace

\vspace{2mm}\hrule

Possible journals to publish: New phytologist, journal of ecology, oikos...

\vspace{3mm}\hrule

*Keywords*: heterospecific pollen, plant reproduction, fitness, interspecific competition, phylogenetic distance. 

\doublespace

\bleft



# INTRODUCTION

**Paragraph 1**

In natural systems plant species normally coexist and share their floral visitors with other species @bascompte2003. This pollinator sharing from the plant perspective at the pre-pollination stage can be negative due to competition @pauw2013 or positive due to facilitation @carvalheiro2014. Once, the floral visitor has arrived to the flower, pollen deposition can take place and with an increasing number of visits the higher the chances of ovule fertilization (Refs).

I like what I have so far BUT we can make the story more complete

Moreover, once the pollinator has landed on the stigma some other issues for the plant species may arise, the arrival of foreign pollen and conspecific pollen loss @morales2008.

**Paragraph 2**

The effect of heterospecific pollen has been widely studied (Morales & Traveset 2008). Invasive species are supposed to have greater negative effect than native ones @arceo2016. Although when non-natives species don´t have greater negative effect we still don´t know why. For this reason, this ecological question is non a native non native one is a trait based issue that is still to be solved. Moreover, the quantity of pollen that integrates in the network can be quite variable ranging from low quatities @bartomeus2008 to intermediate (ref) to high (ref). Moreover, closely related species are supposed to reduce fitness in greater effect but the evidence is scarce and based on independent studies with different methodologies (Arceo-gomez & Ashman 2016) or studies that just check it with a pair of species that are highly related with the aim to understand hybridization costs (refs). There is a need to deepen into how relatednes is involve in the costs of heterospecific pollen effect. Furthermore, following the conceptual trait framework of Ashman and Arceo-Gomez on heterospecific pollen there are good theoretical basis for trait effect. Notwithstanding, non empirical work has tested how really these traits are involved in heterospecific pollen effect.

Explain traits. Put examples

**Paragraph 3**





**Paragraph 4**

The great difficulty of working with pollen in a coflowering community make the understanding of heterospecifc pollen effect a real challenge. For this reason we have created an artificial co-flowering community in a glasshouse to test the effect with all the possible combinations among them. Where we test the folowing hypothesis: 1) Does heterospecific pollen reduce seed set, if so, 2) Does heterospecific pollen effect depend on the relatedness of the species, 3) Does heterospecific pollen effect depend on any floral trait? 

# METHODS

[comment starts]()
Glasshouse trial
•	Species selected and why – how you made them co-flower
•	Give details of sources and planting seeds, growth medium in pots, temperature and light details
•	Hand crosses and how you did them, how you measured seed set over time.
•	Analyses of data – standardization, means, matrices etc.  

•	Analyses and technical difficulties: We calculated effect size by subtracting the mean of the cross pollinated seed set by the mean effect of the HP pollen (explain exactly what figures you used to calculate this) – check with liam about potentially using missing values analyses for the species we don’t have?

Check that the method is working well to prove that your crosses were close to 50% results in SI i.e not all mixes were 50/50% and we have now counted all the pollen to make this a quantitative variable. We also need to factor in the point that we have different total abundances of pollen across our treatments, irrespective of ratios. 
To what extent are differences in the ratios of pollen applied by hand across different plant families influenced by plant traits such as pollen size, morphology and stigma surface type? 

Results – may need to include amount of pollen in models as random factor- prefill matrix with missing value analyse for the species you don’t have.  
Question 1: how do different pollination treatments (100% HP, 50% HP, self and cross ) impact HP pollen across different plant families? Even with 100% HP one (or more species?) still produced seed set. 

Result Effect size of Seed set ~ phylogenetic distance relationship
We found that the variation ?/ mean effect size of seed set is positively related to phylogenetic distance.  This means the more unrelated the species are, the greater the negative impact of heterospecific pollen (give stats  effect size i.e. Procrustes, X = 0.35; P = 0.03)


Question 2 : what are the main traits impacting HP impacts? (compatibility system, pollen size, stigma surface, wet/dry stigma, length of style etc.  

Effect size of seed set ~ floral traits/ reproductive plant traits
We found that the three best terms to explain the variation in seed set is pollen/ovule ratio, stigma width and style length (Stats effect size i.e. X = 0.39, P = 0.02).  
Need to provide correlation matrix for all traits just for 10 species
Show both ways to present this.
Which particular traits do you find significant effects for?
Show this and give stats.
Present plot for each trait and effect size

[comment finishes]()




The study was conducted in a glasshouse at University of New England (Armidale, Australia) from November 2017 to March 2018. Rooms were temperature controlled depending on the requirements of the species with day and night temperature differences. The species selected (Table 1) belonged to three different families, Solanaceae, Brassicaceae and Convolvulaceae. The criteria of species/family selection was based on close/distant related species (see phylogenetic tree for relatedness fig 1), heterogeneous traits, low structural flower complexity and fast life cycle. For the purpose of the experiment all the species where considered as pollen recipient and as pollen donor (see interaction matrix, fig 2). Species were watered once or twice per day and fertilized weekly (NPK 23: 3.95: 14).

Brown and Mitchell 2001 could be a good paper to explain why we pick seed set as a proxy and not fruit set. We cannot see changes on it, losing information with it.

**Hand-pollination**

Foreign pollen effect was studied through two different treatments, one with 50% conspecific pollen and 50% heterospecific pollen and a second one with 100% foreign pollen (N=10). Seed set was the proxy of effect (see Brown and Mitchell 2001, for differences in effect between seed set and fruit set) and "pollen tubes". Moreover, hand cross pollination, hand self pollination, apomixis (bagged emasculated flowers) and natural selfing were tested (N=10). Flowers were emasculated the day prior anthesis and hand pollinated next day with a toothpick. Had-pollination was realized with 3-4 gentle touches on the surface of the stigma. The mixes of pollen were performed on an eppendorf based on the pollen counts maded with Neubaeur chamber (each anther was counted 4 times for 20 different anthers per species).


**Evolutive distance**

Two types of evolutive distances were calculated with MEGA7 thow kinds of markers: 1) Internal transcribed spacer (ITS) and 2) ribulose-bisphosphate carboxylase (RBCL)

**Traits**

Several traits of the ten species were measured. Pollen per anther was counted, number of ovules, stigma width and length and stigmatic area, style width and length, ovary width and length. Moreover stigma type was tested. Self-incompatibility was 





```{r model, echo = FALSE, eval = FALSE}
model <- lm(y ~ x)
```

We used the statistical language `R` `r citep(citation())` for all our analyses. These were implemented in dynamic rmarkdown documents using `knitr` `r citep(citation("knitr"))` and `rmarkdown` `r citep(citation("rmarkdown"))` packages. All the multilevel models were fitted with `lme4` `r citep(citation("lme4"))`. 




# RESULTS





# DISCUSSION
 
Discussion

1.	 What are the implications of the findings?


# CONCLUSIONS



# ACKNOWLEDGEMENTS



# REFERENCES

```{r write_citations, cache=FALSE, include=FALSE}
write.bibtex(file="knitcitations.bib")
```

<div id = "refs"></div>


\eleft

\clearpage


\listoftables


\newpage


\newpage

\clearpage

\listoffigures


\newpage

\newpage

\blandscape

\elandscape

\clearpage

