---
title: 'How phylogenetic relatedness and floral traits are involved in heterospecifc pollen effect in an artificial co-flowering community'

author:  

- name: Jose B. Lanuza
  affilnum: '1,2'
  
- name:  
  affilnum: 3
  email: barragansljose@gmail.com
  
affiliation:

- affilnum: 1
  affil: US Environmental Protection Agency, Office of Research and Development, National Health and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell Drive  Narragansett, RI, 02882, USA
  
- affilnum: 2
  affil: Big Name University, Department of R, City, BN, 01020, USA
  
- affilnum: 3
  affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain
  
output:

  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
    includes:
      in_header: header.tex

  html_document: null
  
  word_document: null

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
abstract: no
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

---


```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")


```


```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```


\singlespace

\vspace{2mm}\hrule

Possible journals to publish: New phytologist, journal of ecology, oikos...

\vspace{3mm}\hrule

*Keywords*: heterospecific pollen, plant reproduction, fitness, competition 

\doublespace

\bleft


# INTRODUCTION

**Paragraph 1**

In natural systems plant species normally coexist and share their floral visitors with other species (Bascompte et al., 2003). This pollinator sharing from the plant perpective can be negative due to competition (refs) or positive due to facilitation (refs). Moreover, once the pollinator has landed on the stigma some other issues for the species fitness may arise, the arrival of foreign pollen and conspecific pollen loss (Morales & Traveset 2008)

**Paragraph 2**

The effect of heterospecific pollen has been widely studied (Morales & Traveset 2008). Native species are supposed to have greater negative effect than native ones @arceo2016. However, the quantity of pollen that integrates in the network can be quite low @bartomeus2008. Moreover, closely related species are supposed to reduce fitness in greater effect but the evidence is scarce and based on independent studies with different methodologies (Arceo-gomez & Ashman 2016) or studies that just check it with a pair of species that are highly related with the aim to understand hybridization costs (refs). There is a need to deepen into how relatednes is involve in the costs of heterospecific pollen effect. Furthermore, following the conceptual trait framework of Ashman and Arceo-Gomez on heterospecific pollen there are good theoretical basis for trait effect. Notwithstanding, non empirical work has tested how really these traits are involved in heterospecific pollen effect.



**Paragraph 3**


**Paragraph 4**

The great difficulty of working with pollen in a coflowering community make the understanding of heterospecifc pollen effect a real challenge. For this reason we have created an artificial co-flowering community in a glasshouse to test the effect with all the possible combinations among them. Where we test the folowing hypothesis: 1) Does heterospecific pollen reduce seed set, if so, 2) Does heterospecific pollen effect depend on the relatedness of the species, 3) Does heterospecific pollen effect depend on any floral trait? 

# METHODS

The study was conducted in a glasshouse at University of New England (Armidale, Australia) from November 2017 to March 2018. Rooms were temperature controlled depending on the requirements of the species with day and night temperature differences. The species selected (Table 1) belonged to three different families, Solanaceae, Brassicaceae and Convolvulaceae. The criteria of species/family selection was based on close/distant related species (see phylogenetic tree for relatedness fig 1), heterogeneous traits, low structural flower complexity and fast life cycle. For the purpose of the experiment all the species where considered as pollen recipient and as pollen donor (see interaction matrix, fig 2). Species were watered once or twice per day and fertilized weekly (NPK 23: 3.95: 14).

Brown and Mitchell 2001 could be a good paper to explain why we pick seed set as a proxy and not fruit set. We cannot see changes on it, losing information with it.

**Hand-pollination**

Foreign pollen effect was studied through two different treatments, one with 50% conspecific pollen and 50% heterospecific pollen and a second one with 100% foreign pollen (N=10). Seed set was the proxy of effect (see Brown and Mitchell 2001, for differences in effect between seed set and fruit set) and "pollen tubes". Moreover, hand cross pollination, hand self pollination, apomixis (bagged emasculated flowers) and natural selfing were tested (N=10). Flowers were emasculated the day prior anthesis and hand pollinated next day with a toothpick. Had-pollination was realized with 3-4 gentle touches on the surface of the stigma. The mixes of pollen were performed on an eppendorf based on the pollen counts maded with Neubaeur chamber (each anther was counted 4 times for 20 different anthers per species).


**Evolutive distance**

Two types of evolutive distances were calculated with MEGA7 thow kinds of markers: 1) Internal transcribed spacer (ITS) and 2) ribulose-bisphosphate carboxylase (RBCL)

**Traits**

Several traits of the ten species were measured. Pollen per anther was counted, number of ovules, stigma width and length and stigmatic area, style width and length, ovary width and length. Moreover stigma type was tested. Self-incompatibility was 





```{r model, echo = FALSE, eval = FALSE}
model <- lm(y ~ x)
```

We used the statistical language `R` `r citep(citation())` for all our analyses. These were implemented in dynamic rmarkdown documents using `knitr` `r citep(citation("knitr"))` and `rmarkdown` `r citep(citation("rmarkdown"))` packages. All the multilevel models were fitted with `lme4` `r citep(citation("lme4"))`. 




# RESULTS





# DISCUSSION

Discuss.



# CONCLUSIONS



# ACKNOWLEDGEMENTS



# REFERENCES

```{r write_citations, cache=FALSE, include=FALSE}
write.bibtex(file="knitcitations.bib")
```

<div id = "refs"></div>


\eleft

\clearpage


\listoftables


\newpage


\newpage

\clearpage

\listoffigures


\newpage

\newpage

\blandscape

\elandscape

\clearpage

