---
title: Working title, The impact of heterospecific pollen on plant reproductive success is mediated by phylogenetic distance and floral reproductive traits 
author:
- affilnum: '1,2,3'
  name: Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader
  email: barragansljose@gmail.com
affiliation:
- affil: US Environmental Protection Agency, Office of Research and Development, National
    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
    27 Tarzwell Drive  Narragansett, RI, 02882, USA
  affilnum: 1
- affil: Big Name University, Department of R, City, BN, 01020, USA
  affilnum: 2
- affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain
  affilnum: 3
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    includes:
      in_header: header.tex
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
  word_document: null
  html_document: null
capsize: normalsize
classoption: a4paper
csl: ecology.csl
documentclass: article
fontsize: 11pt
footerdate: yes
geometry: margin=1in
linenumbers: yes
bibliography:
- references.bib
- knitcitations.bib
spacing: doublespacing
abstract: no
---


```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")


```


```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```


\singlespace

\vspace{2mm}\hrule

Pollinator sharing can have negative consequences for species fitness with the arrival of foreign pollen. However, the costs of heterospecific pollen are not yet well understood. For this reason, we have conducted a glasshouse experiment where we try to understand how phylogenetic relatedness and the different traits of these species are involved in this process. We experimentally crossed 10 species belonging to three different families: Brassicaceae, Solanaceae and Convolvulaceae. Overall, more than 4000 crosses were done and seed set and pollen tubes were considered as proxy of effect. We found that for all species foreign pollen (50% or less) reduced seed set. Moreover, the seed set reduction is not dependent on the degree of relatedness of the pollen donor. However, the effect is governed by the degree of relatedness and the traits of the species recipient. Our results show that the outcome of heterospecific pollen deposition is determined in greater degree by the traits of the pollen recipient than the pollen donor and that certain traits such as compatibility system are crucial to understand the costs of heterospecific pollen.


\vspace{3mm}\hrule

*Keywords*: heterospecific pollen, plant reproduction, fitness, interspecific competition, phylogenetic distance. 

\doublespace

\bleft



# INTRODUCTION

**Paragraph 1**

In natural systems plant species normally coexist and share their floral visitors with other species @bascompte2003. This pollinator sharing from the plant perspective at the pre-pollination stage can be negative due to competition @pauw2013 or positive due to facilitation @carvalheiro2014. Once the floral visitor has arrived to the flower, pollen deposition on the stigma can take place and hence ovule fertilization. An increasing number of visits generally correlates with higher chances of fertilization @engel2003. However this is not always the case, among these possible flower visitors we find also nectar robbers and pollen thiefs @inouye1980 and the quality of pollen that is deposited on the stigma is also highly relevant to the pollination succes @aizen2007. Moreover, other less study issues in the pollination process are conspecific pollen loss and the arrival of foreign pollen which can have important detrimental effects for the species fitness @morales2008 @ashman2013.

**Paragraph 2**

Recent studies have advanced in the ecological importance of heterospecific pollen effect @morales2008 @ashamn2013 @arceo2016. A general overview of heterospecific pollen effect is that it can play an important role on species fitness but seems to be context dependent and not always fitness decreased is produced @morales2008. There are several traits that are postulated to be highly relevant to this process but the correlation with heterospecific pollen effect remains porrly understood. 

Mention relatedness somwhere here.






Moreover, the different methods used make difficult to compare among different experiments. For instance many articles do not describe the foreign pollen quantity used to test for effect or how this pollen was applied (Refs). The small size of pollen, great environmental variability and population outcrossing rates among others make of this field of study a daunting task and higly unpredictable.

**Paragraph 3**


can reduce species fitness (REFS) but seems to be highly contex-dependent. There are hypothesized that some traits can play a crucial role in this species interaction such as stigma type, pollen size, 



Mention invasive species in this paragraph

Few studies have tried to understand how relatedness is involved in the hp effect but generally 


Until our knowledge ....

Rescue from here the useful things:

Invasive species are supposed to have greater negative effect than native ones @arceo2016. Although when non-natives species don´t have greater negative effect we still don´t know why. For this reason, this ecological question is non a native non native one is a trait based issue that is still to be solved. Moreover, the quantity of pollen that integrates in the network can be quite variable ranging from low quatities @bartomeus2008 to intermediate (ref) to high (ref). Moreover, closely related species are supposed to reduce fitness in greater effect but the evidence is scarce and based on independent studies with different methodologies (Arceo-gomez & Ashman 2016) or studies that just check it with a pair of species that are highly related with the aim to understand hybridization costs (refs). There is a need to deepen into how relatednes is involve in the costs of heterospecific pollen effect. Furthermore, following the conceptual trait framework of Ashman and Arceo-Gomez on heterospecific pollen there are good theoretical basis for trait effect. Notwithstanding, non empirical work has tested how really these traits are involved in heterospecific pollen effect.

Explain traits. Put examples

what is closely related? same genus? Just that right, the rest is far related?

I would like to add that the experiments focus on two proxies of effect prezygotic and postzygotic. Why focus on postzygotic? Is the final stage where we can see the effect. Further studies should also study germination rates.

**Paragraph 3**


Traditionally heterospecific pollen effect has focused its attention on different pollen donors as a main driver of different effect. However in this article we want to emphasize that this is true for the cases that the species are higly close related where pollen recognition can take place (eg hybridization) but not when this pollen is from less closely related species which the main driver of effect is determined by the reproductive biology of the female part of the plant(compatibility system, stigma type, stigma area and number of ovules). 


**Paragraph 4**

Sell well our work: We are the first empirical experiment testig the effect of heterospecific pollen with phylogenetic distance

The great difficulty of working with pollen in a coflowering community make the understanding of heterospecifc pollen effect a real challenge. For this reason we have created an artificial co-flowering community in a glasshouse to test the effect with all the possible combinations among them. Where we test the folowing hypothesis: 1) Does heterospecific pollen reduce seed set, if so, 2) Does heterospecific pollen effect depend on the relatedness of the species, 3) Does heterospecific pollen effect depend on any floral trait? 


Maybe another possible hypothesis to test is the reciprocity of the effect of heterospecifc pollen????

Use the sterile species as a proof of the mechanical interference. Was a mistake but seems cool proof!!!


# METHODS

[comment starts]()
Glasshouse trial
•	Species selected and why – how you made them co-flower
•	Give details of sources and planting seeds, growth medium in pots, temperature and light details
•	Hand crosses and how you did them, how you measured seed set over time.
•	Analyses of data – standardization, means, matrices etc.  

•	Analyses and technical difficulties: We calculated effect size by subtracting the mean of the cross pollinated seed set by the mean effect of the HP pollen (explain exactly what figures you used to calculate this) – check with liam about potentially using missing values analyses for the species we don’t have?

Check that the method is working well to prove that your crosses were close to 50% results in SI i.e not all mixes were 50/50% and we have now counted all the pollen to make this a quantitative variable. We also need to factor in the point that we have different total abundances of pollen across our treatments, irrespective of ratios. 
To what extent are differences in the ratios of pollen applied by hand across different plant families influenced by plant traits such as pollen size, morphology and stigma surface type? 

Results – may need to include amount of pollen in models as random factor- prefill matrix with missing value analyse for the species you don’t have.  
Question 1: how do different pollination treatments (100% HP, 50% HP, self and cross ) impact HP pollen across different plant families? Even with 100% HP one (or more species?) still produced seed set. 

Result Effect size of Seed set ~ phylogenetic distance relationship
We found that the variation ?/ mean effect size of seed set is positively related to phylogenetic distance.  This means the more unrelated the species are, the greater the negative impact of heterospecific pollen (give stats  effect size i.e. Procrustes, X = 0.35; P = 0.03)


Question 2 : what are the main traits impacting HP impacts? (compatibility system, pollen size, stigma surface, wet/dry stigma, length of style etc.  

Effect size of seed set ~ floral traits/ reproductive plant traits
We found that the three best terms to explain the variation in seed set is pollen/ovule ratio, stigma width and style length (Stats effect size i.e. X = 0.39, P = 0.02).  
Need to provide correlation matrix for all traits just for 10 species
Show both ways to present this.
Which particular traits do you find significant effects for?
Show this and give stats.
Present plot for each trait and effect size

[comment finishes]()




The study was conducted in a glasshouse at University of New England (Armidale, Australia) from November 2017 to March 2018. Rooms were temperature controlled depending on the requirements of the species with day and night temperature differences. The species selected (Table 1) belonged to three different families, Solanaceae, Brassicaceae and Convolvulaceae. The criteria of species/family selection was based on close/distant related species (see phylogenetic tree for relatedness fig 1), heterogeneous traits, low structural flower complexity and fast life cycle. For the purpose of the experiment all the species where considered as pollen recipient and as pollen donor (see interaction matrix, fig 2). Species were watered once or twice per day and fertilized weekly (NPK 23: 3.95: 14).

Brown and Mitchell 2001 could be a good paper to explain why we pick seed set as a proxy and not fruit set. We cannot see changes on it, losing information with it.

**Hand-pollination**

Foreign pollen effect was studied through two different treatments, one with 50% conspecific pollen and 50% heterospecific pollen and a second one with 100% foreign pollen (N=10). Seed set was the proxy of effect (see Brown and Mitchell 2001, for differences in effect between seed set and fruit set) and "pollen tubes". Moreover, hand cross pollination, hand self pollination, apomixis (bagged emasculated flowers) and natural selfing were tested (N=10). Flowers were emasculated the day prior anthesis and hand pollinated next day with a toothpick. Had-pollination was realized with 3-4 gentle touches on the surface of the stigma. The mixes of pollen were performed on an eppendorf based on the pollen counts maded with Neubaeur chamber (each anther was counted 4 times for 20 different anthers per species).


**Evolutive distance**

Two types of evolutive distances were calculated with MEGA7 thow kinds of markers: 1) Internal transcribed spacer (ITS) and 2) ribulose-bisphosphate carboxylase (RBCL)

**Traits**

Several traits of the ten species were measured. Pollen per anther was counted, number of ovules, stigma width and length and stigmatic area, style width and length, ovary width and length. Moreover stigma type was tested. Self-incompatibility was 





```{r model, echo = FALSE, eval = FALSE}
model <- lm(y ~ x)
```

We used the statistical language `R` `r citep(citation())` for all our analyses. These were implemented in dynamic rmarkdown documents using `knitr` `r citep(citation("knitr"))` and `rmarkdown` `r citep(citation("rmarkdown"))` packages. All the multilevel models were fitted with `lme4` `r citep(citation("lme4"))`. 




# RESULTS





# DISCUSSION
 
Discussion

1.	 What are the implications of the findings?


# CONCLUSIONS



# ACKNOWLEDGEMENTS



# REFERENCES

```{r write_citations, cache=FALSE, include=FALSE}
write.bibtex(file="knitcitations.bib")
```

<div id = "refs"></div>


\eleft

\clearpage


\listoftables


\newpage


\newpage

\clearpage

\listoffigures


\newpage

\newpage

\blandscape

\elandscape

\clearpage

