---
title: working title Compatibility system and stygma size of pollen recipient as main predictors of heterospecific pollen effect
author:
- affilnum: '1,2,3'
  name: Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader
  email: barragansljose@gmail.com
affiliation:
- affil: US Environmental Protection Agency, Office of Research and Development, National
    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
    27 Tarzwell Drive  Narragansett, RI, 02882, USA
  affilnum: 1
- affil: Big Name University, Department of R, City, BN, 01020, USA
  affilnum: 2
- affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain
  affilnum: 3
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    includes:
      in_header: header.tex
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
  word_document: null
  html_document: null
capsize: normalsize
classoption: a4paper
csl: ecology.csl
documentclass: article
fontsize: 11pt
footerdate: yes
geometry: margin=1in
linenumbers: yes
bibliography:
- references.bib
- knitcitations.bib
spacing: doublespacing
abstract: no
---


```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")


```


```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```


\singlespace

\vspace{2mm}\hrule

Pollinator sharing can have negative consequences for species fitness with the arrival of foreign pollen. However, the costs of heterospecific pollen are not yet well understood. For this reason, we have conducted a glasshouse experiment where we try to understand how phylogenetic relatedness and the different traits of these species are involved in this process. We experimentally crossed 10 species belonging to three different families: Brassicaceae, Solanaceae and Convolvulaceae. Overall, more than 4000 crosses were done and seed set and pollen tubes were considered as proxy of effect. We found that for all species foreign pollen (50% or less) reduced seed set. Moreover, the seed set reduction is not dependent on the degree of relatedness of the pollen donor. However, the effect is governed by the degree of relatedness and the traits of the species recipient. Our results show that the outcome of heterospecific pollen deposition is determined in greater degree by the traits of the pollen recipient than the pollen donor and that certain traits such as compatibility system are crucial to understand the costs of heterospecific pollen.


\vspace{3mm}\hrule

*Keywords*: heterospecific pollen, plant reproduction, fitness, interspecific competition, phylogenetic distance. 

\doublespace

\bleft

# INTRODUCTION

**Paragraph 1** General idea to our concept

In natural systems plant species normally coexist and share their floral visitors with other species @bascompte2003. This pollinator sharing from the plant perspective at the pre-pollination stage can be negative due to competition @pauw2013 or positive due to facilitation @carvalheiro2014. Once the floral visitor has arrived to the flower, pollen deposition on the stigma can take place and hence ovule fertilization. An increasing number of visits generally correlates with higher chances of fertilization @engel2003. However this is not always the case, among these possible flower visitors we find also nectar robbers and pollen thiefs @inouye1980 and the quality of pollen that is deposit on the stigma is also highly relevant to the pollination succes @aizen2007. Moreover, other less study issues in the pollination process are conspecific pollen loss and the arrival of foreign pollen which can have important detrimental effects on species fitness @morales2008 @ashman2013.

**Paragraph 2** Introducing topic and knowledge gap

Recent studies have advanced in the ecological understanding of heterospecific pollen effect @morales2008 @ashman2013 @arceo2016. A general overview of foreign pollen arrival is that it can play an important role on species fitness but seems to be context dependent and not always produce a decrease in fitness @morales2008. Part of this unpredictability is due to the enormous variability of foreing pollen transferred in nature, where levels between 0 and 75 percent are seen, but most commonly values ranges between 0 and 20 percent of the total pollen load, being the generalist species the ones that receive greater loads of heterospecific pollen @bartomeus2008 @montgomery2012 @ashman2013 @fang2013. Although heterospecific pollen quantity is fundamental to understand the outcome of the interaction so is the different traits of both pollen donor and recipient. @ashman2013 postulated the first predictive framework for traits of heterospecific pollen effect, where different traits such as compatibility system and pollen size among others seems to be crucial to understand foreing pollen effect. Moreover, in @tong2016 an assymetric effect was shown in a crossing experiment between 6 species of the genus *Pedicularis* where the pollen of long styled species was able to grow the full length of the style on short styled species but not viceversa. Despite these recent caveats, we still lack empirical evidence to affirm what are the main traits that drive heterospecific pollen effect for both pollen donor and recipient at seed production level. Interestingly, to comprehend how these traits interact is also crucial to look at the phylogenetic relatedness of the species. There is a considerable amount of literature of crosses between close related species @brown2001 @arceo2016can @tong2016 but few works focused on heterospecific pollen of far related species which is a more realistic feature in natural systems due to a less likely niche overlap (REF). Although close related species seems to affect greatly species fitness @arceo2016 the importance of heterospecific pollen effect of far and close related species remains to be explored at community level.



**Paragraph 3** Expanding ideas with examples

Traditionally, 

. For this reason, controlled pollination experiments where hand pollination is applied instead of using pollinators as pollen vectors have simplified this task. However, other issues may arise such as overestimation of the effect (REFs), different methods across studies or lack of descriptive methodology for reproducible work (REFs). One of the most common ways to estimate heterospecific pollen effect is through hand pollination with mixes of 50% conspecific pollen and 50% heterospecific pollen (REF). 




I would like to add that the experiments focus on two proxies of effect prezygotic and postzygotic. Why focus on postzygotic? Is the final stage where we can see the effect. Further studies should also study germination rates.

Traditionally heterospecific pollen effect has focused its attention on different pollen donors as a main driver of different effect. However in this article we want to emphasize that this is true for the cases that the species are higly close related where pollen recognition can take place (eg hybridization) but not when this pollen is from less closely related species which the main driver of effect is determined by the reproductive biology of the female part of the plant(compatibility system, stigma type, stigma area and number of ovules). 


**Paragraph 4** Introducing our experiment

The great environmental variability in natural systems and complexity of floral structures make heterospecific pollination studies a daunting task. Although plant-pollinator network and pollen network studies can give a first picture of the importance of foreign pollen is necessary to address how its effect is shaped with both traits and relatedness of the species. For this reason, in this study we have created an artificial co-flowering community with 10 species belonging to three different families where we try to test the following questions: 1) Does heterospecific pollen reduce seed set, if so, 2) Does heterospecific pollen effect depend on the relatedness of the species, 3) Does heterospecific pollen effect depend on any floral trait? 

# METHODS

The study was conducted in a glasshouse at University of New England (Armidale, Australia) from November 2017 to March 2018. Rooms were temperature controlled depending on the requirements of the species with day and night temperature differences. The species selected (Table 1) belonged to three different families, Solanaceae, Brassicaceae and Convolvulaceae. The criteria of species/family selection was based on close/distant related species (see phylogenetic tree for relatedness fig 1), heterogeneous traits, low structural flower complexity and fast life cycle. For the purpose of the experiment all the species where considered as pollen recipient and as pollen donor (see interaction matrix, fig 2). Species were watered once or twice per day and fertilized weekly (NPK 23: 3.95: 14).

Brown and Mitchell 2001 could be a good paper to explain why we pick seed set as a proxy and not fruit set. We cannot see changes on it, losing information with it.

**Hand-pollination**

Foreign pollen effect was studied through two different treatments, one with 50% conspecific pollen and 50% heterospecific pollen and a second one with 100% foreign pollen (N=10). Seed set was the proxy of effect (see Brown and Mitchell 2001, for differences in effect between seed set and fruit set) and "pollen tubes". Moreover, hand cross pollination, hand self pollination, apomixis (bagged emasculated flowers) and natural selfing were tested (N=10). Flowers were emasculated the day prior anthesis and hand pollinated next day with a toothpick. Had-pollination was realized with 3-4 gentle touches on the surface of the stigma. The mixes of pollen were performed on an eppendorf based on the pollen counts maded with Neubaeur chamber (each anther was counted 4 times for 20 different anthers per species).


**Evolutive distance**

Two types of evolutive distances were calculated with MEGA7 for two kinds of markers: 1) Internal transcribed spacer (ITS) and 2) ribulose-bisphosphate carboxylase (RBCL)

**Traits**

The traits measured for each species were pollen per anther, number of ovules, stigma width and length and stigmatic area, style width and length, ovary width and length. Moreover stigma type was tested. Pollen was counted for 20 anthers of each species with 4 replicates per sample with an hemocytometer. Previously anthers were squashed on a known solution with the pippete tip and homogeneize with a vortex for 30 seconds. Ovule number was counted with the help of an stereomicroscope and a small grid over a petri dish from 15 randomly selected flowers. The different morphometrical traits were measured with XXXX. Levels of self incompatibility were estimated by dividing the the fruit set of hand self pollination by hand cross pollination 





```{r model, echo = FALSE, eval = FALSE}
model <- lm(y ~ x)
```

We used the statistical language `R` `r citep(citation())` for all our analyses. These were implemented in dynamic rmarkdown documents using `knitr` `r citep(citation("knitr"))` and `rmarkdown` `r citep(citation("rmarkdown"))` packages. All the multilevel models were fitted with `lme4` `r citep(citation("lme4"))`. 




# RESULTS





# DISCUSSION
 
Discussion

1.	 What are the implications of the findings?


# CONCLUSIONS



# ACKNOWLEDGEMENTS



# REFERENCES

```{r write_citations, cache=FALSE, include=FALSE}
write.bibtex(file="knitcitations.bib")
```

<div id = "refs"></div>


\eleft

\clearpage


\listoftables


\newpage


\newpage

\clearpage

\listoffigures


\newpage

\newpage

\blandscape

\elandscape

\clearpage

