---
title: "Results"
author: "Jose"
date: "13/11/18"
output: pdf_document
#bibliography: references.bib
---

```{r, echo=FALSE, message=FALSE}
library(vegan)
library(xtable)
library(kableExtra)
#load environment
load("myEnvironment_mantel.RData")
```

Now I compare the matrix of effect, which is the matrix of cross scaled seed set minus the matrix of mean scaled seed set with the treatments per species.

#Part 1 Relation between effect of Hp and evolutive distance

1) With the RBCL marker

Mantel effect~evolutive distance
```{r, echo=FALSE, message=FALSE}
mantel_rbcl <- mantel(matrix_scale_effect, evo_distance_rbcl)
mantel_rbcl_sqrt <- mantel(matrix_scale_effect, sqrt(evo_distance_rbcl))
pro_rbcl <- protest(matrix_scale_effect, evo_distance_rbcl)
pro_rbcl_sqrt <- protest(matrix_scale_effect, sqrt(evo_distance_rbcl))

r <- c(as.numeric(mantel_rbcl[3]), as.numeric(mantel_rbcl_sqrt[3]),
       as.numeric(pro_rbcl[6]),as.numeric(pro_rbcl_sqrt[6]))
p <- c(as.numeric(mantel_rbcl[4]),as.numeric(mantel_rbcl_sqrt[4]),
       as.numeric(pro_rbcl[13]),as.numeric(pro_rbcl_sqrt[13]))
Analysis <- c("Mantel", "Mantel", "Procustes", "Procustes")
Type <- c("normal evolutive distance","sqrt evolutive distance","normal evolutive distance", "sqrt evolutive distance")
table<- data.frame(r,p, Analysis, Type)
kable(table, format="markdown")
```

I wonder if the matrices are too small for procustes, is a general trend for all the analysis to have a non significant p value and correlation around these numbers.

2) With the ITS marker
```{r, echo=FALSE, message=FALSE}
mantel_its <- mantel(matrix_scale_effect, evo_distance_its)
mantel_its_sqrt <- mantel(matrix_scale_effect, sqrt(evo_distance_its))
pro_its <- protest(matrix_scale_effect, evo_distance_its)
pro_its_sqrt <- protest(matrix_scale_effect, sqrt(evo_distance_its))

r <- c(as.numeric(mantel_its[3]), as.numeric(mantel_its_sqrt[3]),
       as.numeric(pro_its[6]),as.numeric(pro_its_sqrt[6]))
p <- c(as.numeric(mantel_its[4]),as.numeric(mantel_its_sqrt[4]),
       as.numeric(pro_its[13]),as.numeric(pro_its_sqrt[13]))
Analysis <- c("Mantel", "Mantel", "Procustes", "Procustes")
Type <- c("normal evolutive distance","sqrt evolutive distance","normal evolutive distance", "sqrt evolutive distance")
table<- data.frame(r,p, Analysis, Type)
kable(table, format="markdown")
```


#Part 2 Relation between effect of Hp and traits